>

### 2024-12-10 docs/article.qmd
Revise the formal definition following the cited references

Graph theory is a formal concept depicting connections among objects. A graph consists of nodes and edges, denoted as $\mathcal{G} = (\mathcal{V}, \mathcal{E})$. The nodes form a finite set, $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n$, while the edges $\mathcal{E}$ represent the connection between these nodes. In a directed graph, the edge is an ordered pair of nodes, $(\mathcal{v}_i, \mathcal{v}_j) \in \mathcal{E}$, where the order of $i, j$ matters. Directed graphs may contain self connections $(\mathcal{v}_i, \mathcal{v}_i) \in \mathcal{E}$. In an undirected graph, the edge is an unordered set of nodes $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$, where the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [@estrada2012structure].

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, and $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $a_{ij}$ is the edge: $(\mathcal{v}_i, \mathcal{v}_j) \in \mathcal{E}$ in a directed graph, or $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$ in an undirected graph. In an unweighted graph, $a_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $a_{ij} = 0$ otherwise. In a weighted graph, $a_{ij}$ can take any real value ($a_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

@@

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] is equal to one, a network with similarly connected nodes will have an eigenvector centrality $c_i$ = $\frac{1}{n}$ for each node, where $n$ is the number of nodes. The value of $c_i = \frac{1}{n}$ is the expected eigenvector centrality in a densely connected network where connection $(i, j) \in \mathcal{E}$ from node $\mathcal{V}_i$ to node $\mathcal{V}_j$ has the same probability for every node. Per the central limit theorem, the value $\frac{1}{n}$ can be parameterized as a mean of a normal distribution, where all instances of eigenvector centrality will follow the distribution, $c_i \sim \mathcal{N}(\mu, \sigma)$. A one-sample Student's T-test was performed to assess how much each eigenvector centrality differs from the expected values. As such, eigenvector centrality greater than $\frac{1}{n}$ can be considered high, and the value below it can be considered low. The p-values was corrected for a type-I error using Bonferroni method.

>>>>

Graph theory is a formal concept depicting connections among objects. A graph consists of nodes and edges, denoted as $\mathcal{G} = (\mathcal{V}, \mathcal{E})$. The nodes form a finite set, $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n\}$, while the edges $\mathcal{E}$ represent the connection between these nodes. In a directed graph, the edge is an ordered pair of nodes, $(\mathcal{v}_i, \mathcal{v}_j) \in \mathcal{E}$, where the order of $i, j$ matters. Directed graphs may contain self connections, $(\mathcal{v}_i, \mathcal{v}_i) \in \mathcal{E}$. In an undirected graph, the edge is an unordered set of nodes, $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$, where the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections, $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [@estrada2012structure].

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, where $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $a_{ij}$ is the edge: $(\mathcal{v}_i, \mathcal{v}_j) \in \mathcal{E}$ in a directed graph, or $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$ in an undirected graph. In an unweighted graph, $a_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $a_{ij} = 0$ otherwise. In a weighted graph, $a_{ij}$ can take any real value ($a_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

@@

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] is equal to one, a network with similarly connected nodes will have an eigenvector centrality $c_i$ = $\frac{1}{n}$ for each node, where $n$ is the number of nodes. The value of $c_i = \frac{1}{n}$ is the expected eigenvector centrality in a densely connected network where the occurrence of an edge $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$ has the same probability for all nodes. Per the central limit theorem, the value $\frac{1}{n}$ can be parameterized as a mean of a normal distribution, where all instances of eigenvector centrality will follow the distribution, $c_i \sim \mathcal{N}(\mu, \sigma)$. A one-sample Student's T-test was performed to assess how much each eigenvector centrality differs from the expected values. As such, eigenvector centrality greater than $\frac{1}{n}$ can be considered high, and the value below it can be considered low. The p-values was corrected for a type-I error using Bonferroni method.

### 2024-12-10 docs/exec-summary/_measure-eigen.qmd
Draft an initial report using Elsevier template
Revise the formal definition following the cited references

### Measuring relative importance

As all medications are nodes in a claim network, we can measure the property of each medication using graph theory. One way to represent node property is by calculating node centrality, which can be measured as a number of incoming/outgoing edges (degree) or its connectedness to other nodes (importance). We measured importance using eigenvector centrality, which reflects medication co-occurrence with other important medications. Eigenvector centrality is a recursive measures of how well connected a certain node is to other nodes that is also well connected in the network [@Golbeck2013]. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $\c_i$ and $\c_j$ are the centralities of node $i$ and $j$, respectively; $\lambda$ is the eigenvalue; $a_{j, i}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $a_i$ and $a_j$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} a_{j, i} \cdot c_j
$$ {#eq-eigen-centrality}

>>>>

::: {.content-visible unless-meta="in.article"}
### Measuring relative importance
:::

As all medications are nodes in a claim network, we can measure the property of each medication using graph theory. One way to represent node property is by calculating node centrality, which can be measured as a number of incoming/outgoing edges (degree) or its connectedness to other nodes (importance). We measured importance using eigenvector centrality, which reflects medication co-occurrence with other important medications. Eigenvector centrality is a recursive measures of how well connected a certain node is to other nodes that is also well connected in the network [@Golbeck2013]. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $c_i$ and $c_j$ are the centralities of node $\mathcal{v}_i$ and $\mathcal{v}_j$, respectively; $\lambda$ is the eigenvalue; $a_{ji}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $\mathcal{v}_j$ and $\mathcal{v}_i$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} a_{ji} \cdot c_j
$$ {#eq-eigen-centrality}

### 2024-12-10 docs/exec-summary/_preprocessing.qmd
Incorporate suggestion from Spyros
Revise the formal definition following the cited references

Daily medication claims is a tabular data containing values from single medication entity for each patients in the cohort. Data transformation was performed to construct a square matrix of aggregate medication claim per day, with rows and columns representing groups of medication. Each element in the upper and lower triangle represents co-prescriptions, while the diagonal of the matrix represents single prescriptions. All elements in the matrix is a tally of drug combination/single prescription in the population claimed on a daily basis. The matrix follows a graph construct, where we can denote the row-to-column as `Medication1-PrescribedAlongside-Medication2`. The graph is a representation of drug prescription network (DPN), where the edge is the number of concurrency. In the original DPN, the edge weight was the number of claims by assuming that all of defined daily doses (DDD) were equal to 1 [@Cavallo2012]. However, not all medications were prescribed with DDD equals to 1, and discarding these entries would lead to loss of information. Therefore, we applied weights based on DDD, where the weight is the highest if DDD equals to 1, and gradually diminishes as it is getting further from 1. To adjust the weight, first we set the baseline weight as $\omega_B = 1$. Then, we calculated the weight based on the expected density of a normal distribution, where the expected value is $\mu = \omega_B$ and standard deviation is $\sigma = \frac{\omega_B}{3}$. The expected value was set to $\omega_B$ because we expect entries with DDD equals to 1. The standard deviation was set to $\frac{\omega_B}{3}$ to represent an equidistant deviation from the expected value to 0. As a result, we obtained the weighted DDD as $\omega$.

@@

$$
\stackrel{\textrm{Unregularized}}{
  \begin{bmatrix}
  n_{1, 1} & \dots & n_{1, N} \\
  \vdots & \ddots & \vdots \\
  n_{N, 1} & \dots & n_{N, N} \\
  \end{bmatrix}
}
\to
\stackrel{\textrm{Regularized by } DDD}{
  \begin{bmatrix}
  \frac{1}{2} \cdot \left(\omega_1 + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_1 + \omega_N \right) \\
  \vdots & \ddots & \vdots \\
  \frac{1}{2} \cdot \left(\omega_N + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_N + \omega_N \right) \\
  \end{bmatrix}
}
$$ {#eq-regularize-mtx}

>>>>

Daily medication claims are represented as tabular data where each row corresponds to an individual patient’s prescriptions. The data was transformed into a square matrix summarizing co-prescriptions and single prescriptions to construct a graph-based representation. The rows and columns of the matrix represent groups of medications, while each matrix element indicates the frequency of a specific prescription pattern. Elements in the off-diagonal positions represent co-prescriptions between two medications, while diagonal elements represent single prescriptions. This matrix can be interpreted as a graph, where each medication is a node, and the edges represent co-prescription frequencies. The weight of the edge is derived from the number of daily claims. Originally, all defined daily doses (DDD) were assumed to equal 1, leading to straightforward edge weights based on raw counts. However, since not all medications have DDD values of 1, this assumption could lead to inaccurate representations. To address this, weights were adjusted using the density of a normal distribution. The expected value ($\mu$) was set to 1 to prioritize DDD = 1, and the standard deviation ($\sigma$) was set to $\frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1. This adjustment ensures a more accurate weighting system that reflects real-world prescription patterns while preserving information about medications with non-standard DDD values.

@@

$$
\stackrel{\textrm{Unregularized}}{
  \begin{bmatrix}
  n_{1, 1} & \dots & n_{1, N} \\
  \vdots & \ddots & \vdots \\
  n_{N, 1} & \dots & n_{N, N} \\
  \end{bmatrix}
}
\quad \to \quad
\stackrel{\textrm{Regularized by } DDD}{
  \begin{bmatrix}
  \frac{1}{2} \cdot \left(\omega_1 + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_1 + \omega_N \right) \\
  \vdots & \ddots & \vdots \\
  \frac{1}{2} \cdot \left(\omega_N + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_N + \omega_N \right) \\
  \end{bmatrix}
}
$$ {#eq-regularize-mtx}
